---
title: "DAB501 Final Project"
output:
  pdf_document: default
  html_notebook: default
---

## Student Information

- Name: Kenneth Hudak
- ID: 0752057


## Instructions 

Follow the instructions on the handout provided. 


## Packages and Data

```{r}
library(tidyverse)
library(here)
```

```{r}
# Replace the _____ with the correct file name
df <- read_csv(here('data', 'starbucks_final.csv'))
head(df)
```


## MODELING: First pair of variables

### Question 1
```{r}
ggplot(df, aes(fat, calories)) +
geom_point(color = '#00704A') +
  ggtitle("Exploratory Starbucks *Fat* Content Predicting Calories")
```
#### This scatterplot shows the relationship between Starbucks fat content and the number of calories in a menu item. In this case, fat is on the X axis as the independent variable and calories are on the Y axis as the dependant variable (since it was listed in the assignment as the variable to be predicted.) 

#### Therefore, the explanatory variable is FAT.


### Question 2
The response variable is CALORIES.


### Question 3 - Linear Regression Model

```{r}
fat_lm_model <- lm(calories ~ fat, data=df)
summary(fat_lm_model)
```

### Question 4 - Equations of the Models

Predicted Calories = y-intercept + coefficient * explanatory variable

Fat model equation:
PredictedCalories = 183.734 + 11.267 * fat


### Question 5

The intercept is where the slope line crosses the Y-axis. In this case for the fat variable, it crosses at 183.73. This means that when x=0 then y=183.73. We should expect that Starbucks drinks with 0 fat will be 183.7 calories.


### Question 6

The intersect is useful for this data. It shows that we can predict that even if a drink has 0 fat, it should still have about 183 calories in it. This could be due to sugar content in the drink.


### Question 7

The fat slope in this context means that as fat increases by 1 along the x-asix, then calories along the y-axis should have an associated increase by 11.267.



## MODELING: Second pair of variables

### Question 1

```{r}
ggplot(df, aes(carb, calories)) +
geom_point()+
  ggtitle("Exploratory Starbucks *Carbs* Content Predicting Calories")
```
#### This scatterplot shows the relationship between Starbucks carb content and the number of calories in a menu item. In this case, carb is on the X axis as the independent variable and carb are on the Y axis as the dependant variable (since it was listed in the assignment as the variable to be predicted.) Therefore, the explanatory variable is CARB. 

### Question 2
The response variable is CALORIES.


### Question 3

```{r}
carb_lm_model <- lm(calories ~ carb, data=df)
summary(carb_lm_model)
```

### Question 4 - Equations of the Models

Predicted Calories = y-intercept + coefficient * explanatory variable

Carb model equation
PredictedCalories = 146.0204 + 4.2971 * carb


### Question 5

The intercept is where the slope line crosses the Y-axis. In this case for the CARB variable, it crosses at 146.0204. This means that when x=0 then y=146.02. We should expect that Starbucks drinks with 0 CARBS will be 146.02 calories.



### Question 6

The intersect is useful for this data. It shows that we can predict that even if a drink has 0 carbs, it should still have about 146 calories in it. This could be because the drink is sugar free, but has some fat content in the drink contributing to the 146 calories.


### Question 7

The fat slope in this context means that as carb increases by 1 along the x-axis, then calories along the y-axis should have an associated increase by 4.29.



## MODEL ASSESSMENT

### Question 1

R2 is a good metric for comparing models. The higher the R squared value, the better the model.


### Question 2

According to the class slides, R2 (which is the square of the cor coefficient) shows the strength of the fit of a linear model. For the Starbucks models, r2 will show us what percent of variablity in the response variable is explained by the model. The more the model is explaining the variations the better it is. I'd be looking for a higher R2 to know that the response variable is more closely related to the predictor and not other aspects such as noise or randomness.


### Question 3

According to the r2 metric, the lm() function showed that the fat model has a r2 of 0.5756
The lm() showed carb model to have a r2 of 0.4556.
Since FAT model has a higher r2, it is the better model and explains the amount of fat in a drink accounts for 57.5% of the calorie calculation. 

## MODEL DIAGNOSTICS

### Question 1 - Create 2 new Columns
PredictedCalories = Yintercept + slope * my best model(FAT)
residuals are actual values minus predicted values

```{r}
#get the predicted calories based on fat content
df2 <- df %>% mutate(predictedCalories = 183.734 + 11.267 * fat)

#get the residuals, mutate it into the new df2 table created above
df3 <- df2 %>% mutate(residuals = calories - predictedCalories)
head(df3)

```

### Question 2 - Check Linearity

```{r}
ggplot(df3, aes(x = calories, y = residuals)) +
  geom_point() + 
  geom_hline(yintercept = 0)
```
Linearity is not met. The variablity around the 0-line is not constant (a lack of homoscedasticity). I expected this outcome with only 57% of fat accounting for calories.

### Question 3 - Nearly normal residuals

```{r}
#A histogram will show the distribution

ggplot(df3, aes(x = residuals)) + 
  geom_histogram(fill = '#00704A', colour = 'white', size = 2, binwidth = 50)


```
This histogram is shows it is not a normal distribution. It is unimodal but left skewed because of the -100 bar. Otherwise, it would be very close to a normal distribution without the -100 bar. I would want to see the largest bar be at the zero line.
### Question 4

```{r}
ggplot(df3, aes(x = calories, y = residuals)) +
  geom_point() + 
  geom_hline(yintercept = 0) +
  geom_hline(yintercept = 100, colour = 'green') +
  geom_hline(yintercept = -100, colour = 'red')
```
### Conclusion

### Question 1
The constant variability graph does not have a random variance from the zero line. Too many residuals are falling in a cluster below the -100 red line. This model should not be used to soley determine how many calories a drink has just based on fat alone. We should look for the other 43% of variables that contribute to total calories. Perhaps a new variable could be created to show the ratio of fat and carbs per drink. 





